### 2021-05-10 README.Rmd
Remove packrat

---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
# Please put your title here to include it in the file below.
title_fr <- "Analyse de données de composition et identification des altérations géochimiques des matériaux céramiques : le cas des productions d’un atelier ibérique (Teruel, Espagne ; IIe-Ier siècles avant J.-C.)"
title_en <- "Alterations of ceramic materials: compositional data analysis of an Iberian workshop (Teruel, Spain; 2nd-1st c. BC)"
```

# TaphoCeram

[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.2613512.svg)](https://doi.org/10.5281/zenodo.2613512)
[![GitHub Release](https://img.shields.io/github/release/nfrerebeau/TaphoCeram.svg)](https://github.com/nfrerebeau/TaphoCeram/releases)

This repository contains the data and code for our paper (in french):

> Frerebeau, N., Ben Amara, A. et Cantin, N. (2020). `r title_fr`. _Archéosciences, revue d'Archéométrie_, 44(1), p. 33-50. DOI : <https://doi.org/10.4000/archeosciences.7220>.

The files hosted in this repository are the development versions and may have changed since the paper was published.

## How to cite

Please cite this compendium as:

> Frerebeau, N., Ben Amara, A. and Cantin, N. (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for "`r title_en`"_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/10.5281/zenodo.2613512>.

## Licenses

**Text and figures:** [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/).

**Code:** see the [DESCRIPTION](DESCRIPTION) file.

**Datasets:** [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/).

## How to download or install

You can download the compendium as a zip from from this [URL](http://github.com/nfrerebeau/TaphoCeram/archive/master.zip).

Or you can install this compendium as an R package, `TaphoCeram`, from GitHub with:

```{r gh-installation, eval=FALSE}
# install.packages("remotes")
remotes::install_github("nfrerebeau/TaphoCeram")
```

To download the package source as you see it on GitHub, for offline browsing, use this line at the shell prompt (assuming you have Git installed on your computer):

```{sh, eval=FALSE}
git clone https://github.com/nfrerebeau/TaphoCeram.git
```

## Usage
Once the download is complete, open the `TaphoCeram.Rproj` in RStudio to begin working with the package and compendium files.

The `analysis/` directory contains:

* The Rmarkdown source file of the manuscript (in the `paper/` subdirectory),
* All the figures generated from R code (in the `figures/` subdirectory),
* All the raw data files (in CSV format, in the `data/` subdirectory),
* Code files used for data cleaning (in the `scripts/` subdirectory).

The `R/` directory contains the custom R functions used in this project. Datasets (in RDA format) are located in the `data/` directory.

## Dependencies

The original manuscript was produced with R 3.6.3. See the [DESCRIPTION](DESCRIPTION) file for a full list of the packages that this project depends on, including the suggested packages.